---
title: "PGP - Candidate locations selection"
author: "Eirini Zormpa"
date: "2 May 2024"
format: html
---

## The brief
Identification of **5-10 cities** of interest across at least 2 countries in the UK.
We suggest using the **Indices of Multiple Deprivation** (specific for children if possible) and utilising any other data we already have to inform this (perhaps data from urban futures?).
A PGP partner mentioned **‘Access to nature’** data which may be useful if it’s already in an easily usable format – I’ve asked Eirini to look at it as a secondary focus and any other nature connection indices

### What is a city?

### Index of Multiple Deprivation & Access to Nature
The data presented below comes from the ONS and relates to [Access to gardens and public green space in Great Britain](https://www.ons.gov.uk/economy/environmentalaccounts/datasets/accesstogardensandpublicgreenspaceingreatbritain)


```{r}
#| echo: false

library(here)
library(readxl)
```
This file contains the following sheets


```{r}
#| echo: false
excel_sheets(here("data_raw", "ospublicgreenspacereferencetables.xlsx"))
```







