---
title: "PGP - Candidate locations selection"
author: "Eirini Zormpa"
date: "2 May 2024"
format:
  html:
    theme: flatly
    toc: true
---

<!--
To-do:
- make structure
- do the things needed in each section :D
- make custom RSA theme
-->

## The brief
Identification of **5-10 cities** of interest across at least 2 countries in the UK.
We suggest using the **Indices of Multiple Deprivation** (specific for children if possible) and utilising any other data we already have to inform this (perhaps data from urban futures?).
A PGP partner mentioned **‘Access to nature’** data which may be useful if it’s already in an easily usable format – I’ve asked Eirini to look at it as a secondary focus and any other nature connection indices

## What is a city?
A practical answer to this complicated question comes from the [Centre for Cities](https://www.centreforcities.org): they have selected the UK's 63 largest cities and towns, which they have defined as Primary Output Areas (PUAs).
As explained on their website:

> This is a measure of the “built-up” area of a large city or town, rather than individual local authority districts.
> We use PUAs in our analysis because they provide a consistent measure to compare concentrations of economic activity across the UK.
> This makes them distinct from city region or combined authority geographies.

Following this methodology, the process presented here limits itself to these 63 PUAs.
This excludes more rural areas, which could be problematic.
On the one hand, these places are likely to have better access to green spaces.
On the other hand, these green spaces may not be suitable for children to play in.
Additionally, rural places can suffer from high levels of deprivation.

## Place selection

### Index of Multiple Deprivation & Access to Nature
The data presented below comes from the ONS and relates to [Access to gardens and public green space in Great Britain](https://www.ons.gov.uk/economy/environmentalaccounts/datasets/accesstogardensandpublicgreenspaceingreatbritain).


```{r}
#| echo: true
#| message: false
#| warning: false

library(here)
library(readxl)
library(dplyr)
library(janitor)
library(tidyr)
```

This file contains the following sheets:

```{r}
#| echo: true

excel_sheets(here("data-raw", "ospublicgreenspacereferencetables.xlsx"))
```
To answer the question we're interested in ("Which are the most deprived cities in the UK?") we need to:

1. combine LSOAs into Local Authority Districts (LADs)
  - NB IMD is calculated on the level of LSOA
2. calculate how deprived each LAD is
3. combine LADs into cities/PUAs, as defined by the Centre for Cities
4. calculate how deprived each city/PUA is
5. order cities/PUAs in order of deprivation

Let's start by looking at the data in sheet number nine, *LSOA Parks and Playing Fields*.

```{r}
#| echo: true

lsoa_parks_play <- read_excel(here("data-raw", "ospublicgreenspacereferencetables.xlsx"),
                              sheet = 9) %>% 
  clean_names()

dim(lsoa_parks_play)
names(lsoa_parks_play)
```


```{r}
#| echo: false
puas <- lsoa_parks_play %>% 
  mutate(PUA = case_when(
           lad_name == "Aberdeen" ~ "Aberdeen",
           lad_name == "Rusmoor" ~ "Aldershot",
           lad_name == "Surrey Heath" ~ "Aldershot",
           lad_name == "Barnsley" ~ "Barnsley",
           lad_name == "Basildon" ~ "Basildon",
           lad_name == "Wirral" ~ "Birkenhead",
           lad_name == "Birmingham" ~ "Birmingham",
           lad_name == "Dudley" ~ "Birmingham",
           lad_name == "Sandwell" ~ "Birmingham",
           lad_name == "Solihull" ~ "Birmingham",
           lad_name == "Walsall" ~ "Birmingham",
           lad_name == "Wolverhampton" ~ "Birmingham",
           lad_name == "Blackburn with Darwen" ~ "Blackburn",
           lad_name == "Blackpool" ~ "Blackpool",
           lad_name == "Fylde" ~ "Blackpool",
           lad_name == "Dorset" ~ "Bournemouth",
           lad_name == "Bournemouth, Christchurch and Poole" ~ "Bournemouth",
           lad_name == "Bradford" ~ "Bradford",
           lad_name == "Adur" ~ "Brighton",
           lad_name == "Brighton and Hove" ~ "Brighton",
           lad_name == "City of Bristol" ~ "Bristol",
           lad_name == "South Gloucestershire" ~ "Bristol",
           lad_name == "Burnley" ~ "Burnley",
           lad_name == "Pendle" ~ "Burnley",
           lad_name == "Cambridge" ~ "Cambridge",
           lad_name == "Cardiff" ~ "Cardiff",
           lad_name == "Medway" ~ "Chatham",
           lad_name == "Coventry" ~ "Coventry",
           lad_name == "Crawley" ~ "Crawley",
           lad_name == "Derby" ~ "Derby",
           lad_name == "Doncaster" ~ "Doncaster",
           lad_name == "Dundee" ~ "Dundee",
           lad_name == "Edinburgh" ~ "Edinburgh",
           lad_name == "Exeter" ~ "Exeter",
           lad_name == "East Dunbartonshire" ~ "Glasgow",
           lad_name == "East Renfrewshire" ~ "Glasgow",
           lad_name == "Glasgow" ~ "Glasgow",
           lad_name == "Renfrewshire" ~ "Glasgow",
           lad_name == "Gloucester" ~ "Gloucester",
           lad_name == "Kirklees" ~ "Huddersfield",
           lad_name == "Kingston upon Hull" ~ "Hull",
           lad_name == "Ipswich" ~ "Ipswich",
           lad_name == "Leeds" ~ "Leeds",
           lad_name == "Blaby" ~ "Leicester",
           lad_name == "Leicester" ~ "Leicester",
           lad_name == "Oadby and Wigston" ~ "Leicester",
           lad_name == "Knowsley" ~ "Liverpool",
           lad_name == "Barking and Dagenham" ~ "London",
           lad_name == "Barnet" ~ "London",
           lad_name == "Bexley" ~ "London",
           lad_name == "Brent" ~ "London",
           lad_name == "Bromley" ~ "London",
           lad_name == "Broxbourne" ~ "London",
           lad_name == "Camden" ~ "London",
           lad_name == "City of London" ~ "London",
           lad_name == "Croydon" ~ "London",
           lad_name == "Dartford" ~ "London",
           lad_name == "Ealing" ~ "London",
           lad_name == "Elmbridge" ~ "London",
           lad_name == "Enfield" ~ "London",
           lad_name == "Epping Forest" ~ "London",
           lad_name == "Epsom and Ewell" ~ "London",
           lad_name == "Gravesham" ~ "London",
           lad_name == "Greenwich" ~ "London",
           lad_name == "Hackney" ~ "London",
           lad_name == "Hammersmith and Fulham" ~ "London",
           lad_name == "Haringey" ~ "London",
           lad_name == "Harrow" ~ "London",
           lad_name == "Havering" ~ "London",
           lad_name == "Hertsmere" ~ "London",
           lad_name == "Hillingdon" ~ "London",
           lad_name == "Hounslow" ~ "London",
           lad_name == "Islington" ~ "London",
           lad_name == "Kensington and Chelsea" ~ "London",
           lad_name == "Kingston upon Thames" ~ "London",
           lad_name == "Lambeth" ~ "London",
           lad_name == "Lewisham" ~ "London",
           lad_name == "Merton" ~ "London",
           lad_name == "Newham" ~ "London",
           lad_name == "Redbridge" ~ "London",
           lad_name == "Richard upon Thames" ~ "London",
           lad_name == "Runnymede" ~ "London",
           lad_name == "Southwark" ~ "London",
           lad_name == "Spelthorne" ~ "London",
           lad_name == "Sutton" ~ "London",
           lad_name == "Three Rivers" ~ "London",
           lad_name == "Tower Hamlets" ~ "London",
           lad_name == "Waltham Forest" ~ "London",
           lad_name == "Wandsworth" ~ "London",
           lad_name == "Watford" ~ "London",
           lad_name == "Westminster" ~ "London",
           lad_name == "Woking" ~ "London",
           lad_name == "Luton" ~ "Luton",
           lad_name == "Bolton" ~ "Manchester",
           lad_name == "Bury" ~ "Manchester",
           lad_name == "Manchester" ~ "Manchester",
           lad_name == "Oldham" ~ "Manchester",
           lad_name == "Rochdale" ~ "Manchester",
           lad_name == "Salford" ~ "Manchester",
           lad_name == "Stockport" ~ "Manchester",
           lad_name == "Tameside" ~ "Manchester",
           lad_name == "Trafford" ~ "Manchester",
           lad_name == "Ashfield" ~ "Mansfield",
           lad_name == "Mansfield" ~ "Mansfield",
           lad_name == "Middlesbrough" ~ "Middlesbrough",
           lad_name == "Redcar and Cleveland" ~ "Middlesbrough",
           lad_name == "Stockton-on-Tees" ~ "Middlesbrough",
           lad_name == "Milton Keynes" ~ "Milton Keynes",
           lad_name == "Gateshead" ~ "Newcastle",
           lad_name == "Newcastle upon Tyne" ~ "Newcastle",
           lad_name == "North Tyneside" ~ "Newcastle",
           lad_name == "South Tyneside" ~ "Newcastle",
           lad_name == "Newport" ~ "Newport",
           lad_name == "Torfaen" ~ "Newport",
           lad_name == "West Northamptonshire" ~ "Northampton",
           lad_name == "Broadland" ~ "Norwich",
           lad_name == "Norwich" ~ "Norwich",
           lad_name == "Broxtowe" ~ "Nottingham",
           lad_name == "Erewash" ~ "Nottingham",
           lad_name == "Gedling" ~ "Nottingham",
           lad_name == "Nottingham" ~ "Nottingham",
           lad_name == "Oxford" ~ "Oxford",
           lad_name == "Peterborough" ~ "Peterborough",
           lad_name == "Plymouth" ~ "Plymouth",
           lad_name == "Portsmouth" ~ "Portsmouth",
           lad_name == "Fareham" ~ "Portsmouth",
           lad_name == "Gosport" ~ "Portsmouth",
           lad_name == "Havant" ~ "Portsmouth",
           lad_name == "Chorley" ~ "Preston",
           lad_name == "Preston" ~ "Preston",
           lad_name == "South Ribble" ~ "Preston",
           lad_name == "Reading" ~ "Reading",
           lad_name == "Wokingham" ~ "Reading",
           lad_name == "Rotherham" ~ "Sheffield",
           lad_name == "Sheffield" ~ "Sheffield",
           lad_name == "Slough" ~ "Slough",
           lad_name == "Eastleigh" ~ "Southampton",
           lad_name == "Southampton" ~ "Southampton",
           lad_name == "Castlepoint" ~ "Southend",
           lad_name == "Southend-on-Sea" ~ "Southend",
           lad_name == "Rochford" ~ "Southend",
           lad_name == "Newcastle-under-Lyme" ~ "Stoke",
           lad_name == "Stoke-on-Trent" ~ "Stoke",
           lad_name == "Sunderland" ~ "Sunderland",
           lad_name == "Neath Port Talbot" ~ "Swansea",
           lad_name == "Swansea" ~ "Swansea",
           lad_name == "Swindon" ~ "Swindon",
           lad_name == "Telford and Wrekin" ~ "Telford",
           lad_name == "Wakefield" ~ "Wakefield",
           lad_name == "Warrington" ~ "Warrington",
           lad_name == "Wigan" ~ "Wigan",
           lad_name == "Worthing" ~ "Worthing",
           lad_name == "York" ~ "York"
          ))
```

```{r}
#| echo: false
pua_imd <- puas %>% 
  count(country_name, PUA, index_of_multiple_deprivation_decile_country_specific) %>% 
  group_by(PUA) %>% 
  mutate(percentage = (n/sum(n)) * 100) %>% 
  ungroup() %>%
  filter(index_of_multiple_deprivation_decile_country_specific %in% 1:3) %>% 
  drop_na(PUA)
```



### Cross-checking with harmonised IMD data


### Cross-checking with Childern's IMD (England)




